---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# proclhmm: Latent Hidden Markov Model for Process Data

<!-- badges: start -->
<!-- badges: end -->

This package provides functions for simulating from and fitting the latent hidden Markov models for response process data ([Tang, 2024](https://doi.org/10.1007/s11336-023-09938-1)).
It also includes functions for simulating from and fitting ordinary hidden Markov models.

## Installation

You can install the development version of `proclhmm` from GitHub with

``` {r eval=FALSE}
devtools::install_github("xytangtang/proclhmm")
```

## Basic Usage

### Simulate Parameters and Data

```{r sim}
library(proclhmm)
N <- 10 # number of actions
K <- 3 # number of hidden states

# generate parameters
set.seed(12345)
paras_true <- sim_lhmm_paras(N, K)

n <- 100 # sample size
# generate data
data0 <- sim_lhmm(n, paras_true, min_len = 4, mean_len = 25)
action_seqs <- data0$seqs # action sequences
```


### Estimating parameters and latent traits
```{r est}
# generate initial values of parameters
paras_init <- sim_lhmm_paras(N, K)
# model fitting
lhmm_res <- lhmm(action_seqs, K, paras_init)
# estimated discrimation parameters for state transition probability matrix
lhmm_res$paras_est$para_a
# estimated location parameters for state-action probability matrix
lhmm_res$paras_est$para_beta
```

### Most likely hidden state sequences
```{r hidden_state}
# compute state-transition and state-action probability matrix for the first action sequnece
paras_est <- lhmm_res$paras_est
paras_PQ <- compute_PQ_lhmm(lhmm_res$theta_est[1], paras_est$para_a, paras_est$para_b, paras_est$para_alpha, paras_est$para_beta)
P <- paras_PQ$P
Q <- paras_PQ$Q
# compute initial state probability
P1 <- compute_P1_lhmm(paras_est$para_P1)
# find the most likely hidden state sequences for the first action sequence
find_state_seq(action_seqs[[1]], P1, P, Q)
```

